Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming

### 2020-04-13 Paper/manuscript_v2/manuscript_07_01.tex
Notation

% By marginalizing over $\bm{f}$ and conditioning on vector of observations $\bm{y}$, we obtain the posterior distribution of interest
% %
% \begin{align}\label{eq:posterior}
% 	p(f^*|\bm{y}) = \frac{\int p(\bm{y}|\bm{f})p(\bm{f}, f^*)\text{d}\bm{f}}{p(\bm{y})},
% \end{align}
% where $p(\bm{y})$ is the marginal likelihood and is given by
% %
% \begin{align}\label{eq:evidence}
% 	p(\textbf{y}) = \int p(\bm{y}|\bm{f})p(\bm{f}, f^*) \text{d}\bm{f}\text{d}f^*.
% \end{align}
% %
% If the observation model $p(\bm{y}|\bm{f})$ is Gaussian, both integrals in eq. \eqref{eq:posterior} and \eqref{eq:evidence} can be solved analytically conditioned on the hyperparameters. For example, an isotropic Gaussian likelihood yields the following closed-form solution
% \begin{align*}
% 	p(f^*| \bm{y}) = \mathcal{N}(f^*| \textbf{k}_{f^*,\bm{f}} (\textbf{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{f},  k_{f^*,f^*}-\bm{k}_{f^*,\bm{f}} (\bm{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{k}_{\bm{f},f^*}  ),
% \end{align*}
% where $\sigma^2$ is the noise variance.

@@

Thus, the Gram matrix $\mathrm{K}$ of the covariance function $k$ for a set of observations $i=1,\ldots,n$ and corresponding input values $\{x_i\}_{i=1}^{n} \in \Omega^{n}$ can be represented as
%
\begin{equation}
\mathrm{K} = \Phi \Delta \Phi^\intercal, \nonumber
\end{equation}
where $\Phi \in {\rm I\!R}^{n\times m}$ is the matrix of eigenfunctions $\phi_j(x_i)$
%
\begin{align}
\Phi =  \left[ {\begin{array}{ccc}
   \phi_1(x_1) & \cdots & \phi_m(x_1)  \\
    \vdots &\ddots & \vdots  \nonumber \\ 
    \phi_1(x_n) & \cdots & \phi_m(x_n) \\
  \end{array} } \right].
\end{align}
As a result, the model for $f$ can be written as
%
\begin{equation}
\bm{f} \sim \mathcal{N}(\bm{\mu},\Phi \Delta \Phi^\intercal). \nonumber
\end{equation}
This equivalently leads to a linear representation of $f$ via
%
\begin{equation}\label{eq_approxf}
f(x) \approx \sum_{j}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{\frac{1}{2}} \phi_j(x) \beta_j,
\end{equation}
%
where $\beta_j \sim \mathcal{N}(0,1)$. Thus, the function $f$ is approximated with a finite basis function expansion (using the eigenfunctions $\phi_j$ of the Laplace operator), scaled by the square root of spectral density values. A key property of this approximation is that the eigenfunctions $\phi_j$ do not depend on the hyperparameters of the covariance function $\theta$. Instead, the only dependence of the model on $\theta$ is through the spectral density $S_{\theta}$. The eigenvalues $\lambda_j$ are monotonically increasing with $j$ and $S_{\theta}$ goes rapidly to zero for bounded covariance functions. Therefore, eq.~\eqref{eq_approxf} can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$. The computational cost of evaluating the log posterior density of univariate HSGPs scales as $O(nm + m)$, where $n$ is the number of observations and $m$ the number of basis functions.

@@

As explained in Section~\ref{ch5_sec_method}, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset {\rm I\!R}$
%
\begin{equation*}
k(\tau) = \sum_{j=1}^{\infty} S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0), 
\end{equation*}
%
where $L$ describes the boundary condition, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$ is the difference between two covariate values $x$ and $x'$ in $\Omega$. The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}) for the unidimensional case and in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}) for the multidimensional case. The number of basis functions can be truncated at some finite positive value $m$ such that the total variation difference between the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$:
%
\begin{equation}\label{eq_diff_covs}
 \int | k(\tau) - \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0)|  \mathrm{d}\tau  < \varepsilon.
\end{equation}

@@

The component $f_2(t)$ represents the yearly smooth seasonal pattern, using a periodic squared exponential covariance function (with period 365.25 to match the average length of the year) in a GP model,
%
\begin{equation*}
f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\left(\!-\frac{2\,\mathrm{sin}^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
\end{equation*}
% which means the function values $\bm{f}_2=\{f_2(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_2$, where $K_{2_{t,s}}=k_2(t,s)$, with $t,s=1,\dots,T$.

The component $f_3(t)$ represents the weekly smooth pattern using a periodic squared exponential covariance function (with period 7 of length of the week) in a GP model,
%
\begin{equation*}
f_3(t) \sim \GP(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\!\!\left(\!-\frac{2\,\text{sin}^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
\end{equation*}
% which means the function values $\bm{f}_3=\{f_3(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_3$, where $K_{3_{t,s}}=k_3(t,s)$, with $t,s=1,\dots,T$.

@@

We do not have a full observation model, as we do not have a model for the censoring process. We use the complementary cumulative log-normal probability distribution for the censored data conditionally on the censoring time $t_i$:
%
\begin{equation*}
p(y_i > t_i)= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \mathrm{d}y_i=  1 - \Phi\! \left( \frac{\mathrm{log}(y_i)-f(\bm{x}_i)}{\sigma} \right),
\end{equation*}
%
where $y_i>t_i$ denotes the unobserved survival time.

@@

Associated to each covariance function $k(\bm{x},\bm{x}')$ we can also define a covariance operator $\mathcal{K}$ over a function $f(\bm{x})$ as follows:
%
\begin{equation*}
\mathcal{K} f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') \mathrm{d}\bm{x}'.
\end{equation*}

From the Bochner’s and Wiener-Khintchine theorems, the spectral density of a stationary covariance function $k(\bm{x},\bm{x}') = k(\bm{\tau})$, $\bm{\tau}=(\bm{x}-\bm{x}')$, is the Fourier transform of the covariance function,
%
\begin{equation*}
S(\bm{w}) = \int k(\bm{\tau}) e^{-2\pi i \bm{w} \bm{\tau}} \mathrm{d}\bm{\tau}, \nonumber
\end{equation*}
where $\bm{w}$ is in the frequency domain. The operator $\mathcal{K}$ will be translation invariant if the covariance function is stationary. This allows for a Fourier representation of the operator $\mathcal{K}$ as a transfer function which is the spectral density of the Gaussian process. 
%When the covariance function is homogeneous, the corresponding operator $\mathcal{K}$ will be translation invariant thus allowing for Fourier-representation as a transfer function. This transfer function is the spectral density of the Gaussian process
%%
%\begin{eqnarray*}
%S(\bm{w}) = \mathcal{F}(\mathcal{K}). \nonumber
%\end{eqnarray*}
%
%\noindent 
Thus, the spectral density $S(\bm{w})$ also gives the approximate eigenvalues of the operator $\mathcal{K}$.

@@

If the negative Laplace operator $-\nabla^2$ is defined as the covariance operator of the formal kernel $l$,
%
\begin{equation*}
-\nabla^2 f(\bm{x}) = \int l(\bm{x},\bm{x}') f(\bm{x}') \mathrm{d}\bm{x}',
\end{equation*} 
then the formal kernel can be represented as 
%
\begin{equation*}
l(\bm{x},\bm{x}')= \sum_j \lambda_j \phi_j(\bm{x}) \phi_j(\bm{x}'),
\end{equation*}
where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(\bm{x})\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator. Namely, they satisfy the following eigenvalue problem in the compact subset $\bm{x} \in \Omega \subset {\rm I\!R}^D$ and with the Dirichlet boundary condition (other boundary conditions could be used as well):
%
\begin{align*}
-\nabla^2 \phi_j(\bm{x})&=\lambda \phi_j(\bm{x}), \hspace{1cm}  x\in \Omega \nonumber \\ 
\phi_j(\bm{x})&=0, \hspace{1.85cm} x\notin \Omega.
\end{align*}  
Because $-\nabla^2$ is a positive definite Hermitian operator, the set of eigenfunctions $\phi_j(\cdot)$ are orthonormal with respect to the inner product
%
\begin{align*}
<f,g>=\int_\Omega f(\bm{x}) g(\bm{x}) \mathrm{d}(\bm{x})
\end{align*} 
that is,
%
\begin{align*}
\int_\Omega \phi_i(\bm{x}) \phi_j(\bm{x}) \mathrm{d}(\bm{x}) = \delta_{ij},
\end{align*} 
and all the eigenvalues $\lambda_j$ are real and positive.

Due to normality of the basis of the representation of the formal kernel $l(\bm{x},\bm{x}')$, its formal powers $s=1,2,\dots$ can be written as
%
\begin{eqnarray}\label{ch5_eq_formalkernel}
l(\bm{x},\bm{x}')^s= \sum_j \lambda_j^s \phi_j(\bm{x}) \phi_j(\bm{x}'),
\end{eqnarray} 
which are again to be interpreted to mean that
%
\begin{equation*}
(-\nabla^2)^s f(\bm{x}) = \int l^s(\bm{x},\bm{x}') f(\bm{x}') \mathrm{d}\bm{x}'.
\end{equation*} 
This implies that we also have
%
\begin{equation*}
[a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+\cdots] f(\bm{x}) = \int [a_0+a_1l^1(\bm{x},\bm{x}')+a_2l^2(\bm{x},\bm{x}')+\cdots] f(\bm{x}')  \mathrm{d}\bm{x}'.
\end{equation*}

@@

%If the requirement of a valid covariance function is relaxed and only an optimal series approximation is required, taking the limit $J \rightarrow \infty$ in the sub-sums in equation (\ref{ch5_eq_q}) gives the following variance coefficients
An upper bounded approximation to the coefficients $\tilde{q}_j^2$ and $\tilde{q}_0^2$ can be obtained by taking the limit $J \rightarrow \infty$ in the sub-sums in the corresponding equations (\ref{ch5_eq_q}) and (\ref{ch5_eq_q0}), and thus leading to the following variance coefficients:
%
\begin{equation}\label{ch5_eq_q_2}
\begin{split}
\tilde{q}_j^2= \frac{2\text{I}_j(\ell^{-2})}{\exp(\frac{1}{\ell^2})}, \\
\tilde{q}_0^2= \frac{\text{I}_0(\ell^{-2})}{\exp(\frac{1}{\ell^2})},
\end{split}
\end{equation} 
for $j=1,2,\cdots,J$, and where the $\text{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind. This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required. A more detailed explanation and mathematical proofs of this approximation of a periodic covariance function are provided by \citet{solin2014explicit}.

>>>>

% By marginalizing over $\bm{f}$ and conditioning on vector of observations $\bm{y}$, we obtain the posterior distribution of interest
% %
% \begin{align}\label{eq:posterior}
% 	p(f^*|\bm{y}) = \frac{\int p(\bm{y}|\bm{f})p(\bm{f}, f^*)\mathrm{d}\bm{f}}{p(\bm{y})},
% \end{align}
% where $p(\bm{y})$ is the marginal likelihood and is given by
% %
% \begin{align}\label{eq:evidence}
% 	p(\textbf{y}) = \int p(\bm{y}|\bm{f})p(\bm{f}, f^*) \mathrm{d}\bm{f}\mathrm{d}f^*.
% \end{align}
% %
% If the observation model $p(\bm{y}|\bm{f})$ is Gaussian, both integrals in eq. \eqref{eq:posterior} and \eqref{eq:evidence} can be solved analytically conditioned on the hyperparameters. For example, an isotropic Gaussian likelihood yields the following closed-form solution
% \begin{align*}
% 	p(f^*| \bm{y}) = \mathcal{N}(f^*| \textbf{k}_{f^*,\bm{f}} (\textbf{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{f},  k_{f^*,f^*}-\bm{k}_{f^*,\bm{f}} (\bm{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{k}_{\bm{f},f^*}  ),
% \end{align*}
% where $\sigma^2$ is the noise variance.

@@

Thus, the Gram matrix $K$ of the covariance function $k$ for a set of observations $i=1,\ldots,n$ and corresponding input values $\{x_i\}_{i=1}^{n} \in \Omega^{n}$ can be represented as
%
\begin{equation}
  K = \Phi \Delta \Phi^\intercal, \nonumber
\end{equation}
%
where $\Phi \in {\rm I\!R}^{n\times m}$ is the matrix of eigenfunctions $\phi_j(x_i)$
%
\begin{align}
\Phi =  \left[ {\begin{array}{ccc}
   \phi_1(x_1) & \cdots & \phi_m(x_1)  \\
    \vdots &\ddots & \vdots  \nonumber \\ 
    \phi_1(x_n) & \cdots & \phi_m(x_n) \\
  \end{array} } \right].
\end{align}
As a result, the model for $f$ can be written as
%
\begin{equation}
\bm{f} \sim \mathcal{N}(\bm{\mu},\Phi \Delta \Phi^\intercal). \nonumber
\end{equation}
This equivalently leads to a linear representation of $f$ via
%
\begin{equation}\label{eq_approxf}
f(x) \approx \sum_{j}^m \left( S_{\theta}(\sqrt{\lambda_j})\right)^{\frac{1}{2}} \phi_j(x) \beta_j,
\end{equation}
%
where $\beta_j \sim \mathcal{N}(0,1)$. Thus, the function $f$ is approximated with a finite basis function expansion (using the eigenfunctions $\phi_j$ of the Laplace operator), scaled by the square root of spectral density values. A key property of this approximation is that the eigenfunctions $\phi_j$ do not depend on the hyperparameters of the covariance function $\theta$. Instead, the only dependence of the model on $\theta$ is through the spectral density $S_{\theta}$. The eigenvalues $\lambda_j$ are monotonically increasing with $j$ and $S_{\theta}$ goes rapidly to zero for bounded covariance functions. Therefore, eq.~\eqref{eq_approxf} can be expected to be a good approximation for a finite number of $m$ terms in the series as long as the inputs values $x_i$ are not too close to the boundaries $-L$ and $L$ of $\Omega$. The computational cost of evaluating the log posterior density of univariate HSGPs scales as $O(nm + m)$, where $n$ is the number of observations and $m$ the number of basis functions.

@@

As explained in Section~\ref{ch5_sec_method}, the approximation of the covariance function is a series expansion of eigenfunctions and eigenvalues of the Laplace operator in a given domain $\Omega$, for instance in a one-dimensional input space $\Omega=[-L,L]\subset {\rm I\!R}$
%
\begin{equation*}
k(\tau) = \sum_{j=1}^{\infty} S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0), 
\end{equation*}
%
where $L$ describes the boundary condition, $j$ is the index for the eigenfunctions and eigenvalues, and $\tau=x-x'$ is the difference between two covariate values $x$ and $x'$ in $\Omega$. The eigenvalues $\lambda_j$ and eigenfunctions $\phi_j$ are given in equations (\ref{eq_eigenvalue}) and (\ref{eq_eigenfunction}) for the unidimensional case and in equations (\ref{eq_eigenvalue_multi}) and (\ref{eq_eigenfunction_multi}) for the multidimensional case. The number of basis functions can be truncated at some finite positive value $m$ such that the total variation difference between the exact and approximate covariance functions is less than a predefined threshold $\varepsilon > 0$:
%
\begin{equation}\label{eq_diff_covs}
 \int | k(\tau) - \sum_{j=1}^m S_{\theta}(\sqrt{\lambda_j}) \phi_j(\tau) \phi_j(0)|  \,\mathrm{d}\tau  < \varepsilon.
\end{equation}

@@

The component $f_2(t)$ represents the yearly smooth seasonal pattern, using a periodic squared exponential covariance function (with period 365.25 to match the average length of the year) in a GP model,
%
\begin{equation*}
f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \sigma_2^2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
\end{equation*}
% which means the function values $\bm{f}_2=\{f_2(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_2$, where $K_{2_{t,s}}=k_2(t,s)$, with $t,s=1,\dots,T$.

The component $f_3(t)$ represents the weekly smooth pattern using a periodic squared exponential covariance function (with period 7 of length of the week) in a GP model,
%
\begin{equation*}
f_3(t) \sim \GP(0,k_3), \hspace{5mm} k_3(t,t') = \sigma_3^2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
\end{equation*}
% which means the function values $\bm{f}_3=\{f_3(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_3$, where $K_{3_{t,s}}=k_3(t,s)$, with $t,s=1,\dots,T$.

@@

We do not have a full observation model, as we do not have a model for the censoring process. We use the complementary cumulative log-normal probability distribution for the censored data conditionally on the censoring time $t_i$:
%
\begin{equation*}
p(y_i > t_i)= \int_{t_i}^{\infty} \LogNormal(y_i \mid f(\bm{x}_i),\sigma^2) \,\mathrm{d}y_i=  1 - \Phi\! \left( \frac{\log(y_i)-f(\bm{x}_i)}{\sigma} \right),
\end{equation*}
%
where $y_i>t_i$ denotes the unobserved survival time.

@@

Associated to each covariance function $k(\bm{x},\bm{x}')$ we can also define a covariance operator $\mathcal{K}$ over a function $f(\bm{x})$ as follows:
%
\begin{equation*}
\mathcal{K} f(\bm{x}) = \int k(\bm{x},\bm{x}') f(\bm{x}') \,\mathrm{d}\bm{x}'.
\end{equation*}

From the Bochner’s and Wiener-Khintchine theorems, the spectral density of a stationary covariance function $k(\bm{x},\bm{x}') = k(\bm{\tau})$, $\bm{\tau}=(\bm{x}-\bm{x}')$, is the Fourier transform of the covariance function,
%
\begin{equation*}
S(\bm{w}) = \int k(\bm{\tau}) e^{-2\pi i \bm{w} \bm{\tau}} \,\mathrm{d}\bm{\tau}, \nonumber
\end{equation*}
where $\bm{w}$ is in the frequency domain. The operator $\mathcal{K}$ will be translation invariant if the covariance function is stationary. This allows for a Fourier representation of the operator $\mathcal{K}$ as a transfer function which is the spectral density of the Gaussian process. 
%When the covariance function is homogeneous, the corresponding operator $\mathcal{K}$ will be translation invariant thus allowing for Fourier-representation as a transfer function. This transfer function is the spectral density of the Gaussian process
%%
%\begin{eqnarray*}
%S(\bm{w}) = \mathcal{F}(\mathcal{K}). \nonumber
%\end{eqnarray*}
%
%\noindent 
Thus, the spectral density $S(\bm{w})$ also gives the approximate eigenvalues of the operator $\mathcal{K}$.

@@

If the negative Laplace operator $-\nabla^2$ is defined as the covariance operator of the formal kernel $l$,
%
\begin{equation*}
-\nabla^2 f(\bm{x}) = \int l(\bm{x},\bm{x}') f(\bm{x}') \,\mathrm{d}\bm{x}',
\end{equation*} 
then the formal kernel can be represented as 
%
\begin{equation*}
l(\bm{x},\bm{x}')= \sum_j \lambda_j \phi_j(\bm{x}) \phi_j(\bm{x}'),
\end{equation*}
where $\{\lambda_j\}_{j=1}^{\infty}$ and $\{\phi_j(\bm{x})\}_{j=1}^{\infty}$ are the set of eigenvalues and eigenvectors, respectively, of the Laplacian operator. Namely, they satisfy the following eigenvalue problem in the compact subset $\bm{x} \in \Omega \subset {\rm I\!R}^D$ and with the Dirichlet boundary condition (other boundary conditions could be used as well):
%
\begin{align*}
-\nabla^2 \phi_j(\bm{x})&=\lambda \phi_j(\bm{x}), \hspace{1cm}  x\in \Omega \nonumber \\ 
\phi_j(\bm{x})&=0, \hspace{1.85cm} x\notin \Omega.
\end{align*}  
Because $-\nabla^2$ is a positive definite Hermitian operator, the set of eigenfunctions $\phi_j(\cdot)$ are orthonormal with respect to the inner product
%
\begin{align*}
<f,g>=\int_\Omega f(\bm{x}) g(\bm{x}) \,\mathrm{d}(\bm{x})
\end{align*} 
that is,
%
\begin{align*}
\int_\Omega \phi_i(\bm{x}) \phi_j(\bm{x}) \,\mathrm{d}(\bm{x}) = \delta_{ij},
\end{align*} 
and all the eigenvalues $\lambda_j$ are real and positive.

Due to normality of the basis of the representation of the formal kernel $l(\bm{x},\bm{x}')$, its formal powers $s=1,2,\dots$ can be written as
%
\begin{eqnarray}\label{ch5_eq_formalkernel}
l(\bm{x},\bm{x}')^s= \sum_j \lambda_j^s \phi_j(\bm{x}) \phi_j(\bm{x}'),
\end{eqnarray} 
which are again to be interpreted to mean that
%
\begin{equation*}
(-\nabla^2)^s f(\bm{x}) = \int l^s(\bm{x},\bm{x}') f(\bm{x}') \,\mathrm{d}\bm{x}'.
\end{equation*} 
This implies that we also have
%
\begin{equation*}
[a_0+a_1(-\nabla^2)+a_2(-\nabla^2)^2+\cdots] f(\bm{x}) = \int [a_0+a_1l^1(\bm{x},\bm{x}')+a_2l^2(\bm{x},\bm{x}')+\cdots] f(\bm{x}')  \,\mathrm{d}\bm{x}'.
\end{equation*}

@@

%If the requirement of a valid covariance function is relaxed and only an optimal series approximation is required, taking the limit $J \rightarrow \infty$ in the sub-sums in equation (\ref{ch5_eq_q}) gives the following variance coefficients
An upper bounded approximation to the coefficients $\tilde{q}_j^2$ and $\tilde{q}_0^2$ can be obtained by taking the limit $J \rightarrow \infty$ in the sub-sums in the corresponding equations (\ref{ch5_eq_q}) and (\ref{ch5_eq_q0}), and thus leading to the following variance coefficients:
%
\begin{equation}\label{ch5_eq_q_2}
\begin{split}
\tilde{q}_j^2= \frac{2\mathrm{I}_j(\ell^{-2})}{\exp(\frac{1}{\ell^2})}, \\
\tilde{q}_0^2= \frac{\mathrm{I}_0(\ell^{-2})}{\exp(\frac{1}{\ell^2})},
\end{split}
\end{equation} 
for $j=1,2,\cdots,J$, and where the $\mathrm{I}_{j}(z)$ is the modified Bessel function \citep{handbook1970m} of the first kind. This approximation implies that the requirement of a valid covariance function is relaxed and only an optimal series approximation is required. A more detailed explanation and mathematical proofs of this approximation of a periodic covariance function are provided by \citet{solin2014explicit}.
